Evidence of widspread maintenence of functional composition in vertebrate communities

### 2021-09-08 paper/manuscript.Rmd
Flesh out citations

While functional loss is frequently cited as one of the most pressing concerns of the anthropocene [@cardinale2012; @dirzo2014; @young2016], it is not necessarily inevitable even in scenarios of species loss [@daz2001]. Forecasts of functional loss range from negligible [@gallagher2013] to dire [@petchey2002; @pimiento2020]. And while some observed trends show significant functional loss [@flynn2009] others document no loss even in some of the most heavily impacted communities [@edwards2013; @matuoka2020]. On paleoecological time scales, functional composition is often conserved alongside species diversity even in the face of drastic environmental change [@jackson2015; @barnosky2004].
It is therefore critical to establish whether or not functional loss is the general rule for communities on time scales relevant for current environmental impacts.

Assessments of broad-scale temporal change in functional diversity have previously been limited by a lack of functional trait data.
<!-- and computational power? --> The majority of work has therefore focused largely on system-specific studies with traits collected in situ.
Ongoing efforts to assemble functional traits for a variety of taxa have made synthesis of existing community assemblage data and functional traits possible for the first time, providing initial insights into the ways functional diversity changes on a broad scale for specific taxa [e.g. fish @trindade-santos2020, birds @jarzyna2016; @barnagaud2017].
<!-- so what do these studies say? does that go here? --> However, to date there has been no cross-taxa assessment of temporal functional change for a broad geographic and taxonomic extent.

Here we perform the first multi-taxa, multi-system assessment of functional diversity change through time.
We assess thousands of mammal, bird, and amphibian functional diversity timeseries to 1) establish a general trend of functional change 2) ???.

@@

Communities demonstrated almost universal maintenance of functional composition.
While the majority of the studies (\~70%) included in our data experienced significant species turnover, only three experienced a significant shift in any functional dimension.
This suggests certain characteristics of the functional space are conserved even in the face of significant change in species identity, specifically the size of the functional space occupied by the community (FRic) and the distribution of species and individuals within that space (FEve and FDiv).
On average, species additions have similar functional characteristics as lost species and therefore maintain the functional space structure.
<!-- While we did not explicitly assess how turnover rates compare to background turnover, --> <!-- language to distinguish between "functional traits" and a species' role in the context of the functional space -->

-   Paragraph about analog communities?

-   Paragraph about implications for BEF work

Our work challenges assumptions that functional loss is the default state of all or even many communities.
While we do not directly assess the disturbance histories of included communities, many are from sites that experienced significant disturbance or manual treatment (see appendix for metadata of included studies).
<!-- check this, get numbers? --> This is good news for ecosystem function!
But we may also be measuring the wrong dimensions of the trait space or the wrong traits.
Can't really say BEF needs to measure function directly because these communities aren't losing species or functional diversity - turnover is either within regular bounds or if not (if more turnover than normal) it's still maintaining community structure.
And presuming that the original structure was good for function, maintaining it is also good for function!

@@

This study represents the best available data and biodiversity synthesis approaches.
Why then do we find no change in functional diversity despite extensive predictions to the contrary from both **theory** <!-- need more theory in the intro? --> and and many local studies?

@@

\<!-
- You'd imagine the pattern would still be visible even with North American birds...

To further explore the issue we look for example at the North American Breeding Birdy Survey, included in this study and assessed in many other publications.
--\>

>>>>

While functional loss is frequently cited as one of the most pressing concerns of the anthropocene [@cardinale2012; @dirzo2014; @young2016], it is not necessarily inevitable even in scenarios of species loss [@daz2001]. Forecasts of functional loss range from negligible [@gallagher2013] to dire [@petchey2002; @pimiento2020]. And while some observed trends show significant functional loss [@flynn2009] others document no loss even in some of the most heavily impacted communities [@edwards2013; @matuoka2020]. On paleoecological time scales, functional composition is often conserved alongside species diversity even in the face of drastic environmental change [@barnosky2004 @jackson2015] .
It is therefore critical to establish whether or not functional loss is the general rule for communities on time scales relevant for environmental impacts and management action.

Assessments of broad-scale temporal change in functional diversity have previously been limited by a lack of functional trait data.
<!-- and computational power? --> The majority of work has therefore focused largely on system-specific studies with traits collected in situ.
Ongoing efforts to assemble functional traits for a variety of taxa have made synthesis of existing community assemblage data and functional traits possible for the first time, providing initial insights into the ways functional diversity changes on a broad scale for specific taxa [e.g. fish @trindade-santos2020, birds @jarzyna2016; @barnagaud2017].
However, to date there has been no cross-taxa assessment of temporal functional change for a broad geographic and taxonomic extent.

Here we perform the first multi-taxa, multi-system assessment of functional diversity change through time.
We assess thousands of mammal, bird, and amphibian functional diversity timeseries to assess if systematic functional shifts are occurring and to classify the nature of those shifts.

@@

Communities demonstrated almost universal maintenance of functional composition.
While the majority of the studies (\~70%) included in our data experienced significant species turnover, only three experienced a significant shift in any functional dimension.
This suggests certain characteristics of the functional space are conserved even in the face of significant change in species identity, specifically the size of the functional space occupied by the community (FRic) and the distribution of species and individuals within that space (FEve and FDiv).
On average, species additions have similar functional characteristics as lost species and therefore maintain the structure of the functional space.
<!-- While we did not explicitly assess how turnover rates compare to background turnover, --> Our results give strong evidence that maintenance of functional structure is not just a phenomena of paleoecological time scales but also happens on annual time scales [@barnosky2004; @jackson2015].

Our work challenges assumptions that functional loss is the default state of all or even many communities.
While we do not directly assess the disturbance histories of included communities, trends were consistent even for the longest running and most heavily impacted studies.
The North American Breeding Survey for example is made up exclusively of roadside surveys and is therefore in relatively close proximity to human impact [@pardieck2020] . It is considered an authoritative dataset on the state of bird populations on the continent and underpins policy decisions about bird conservation [@rosenberg2019; @sauer2017].
No more robust dataset exists to capture North American avian community change, yet we detected no general shifts in functional structure across the dataset.
Further, none of the 5 included studies that experienced a manual manipulation (e.g. burning, grazing exclosure, etc) experienced any significant functional trends.

<!-- Widespread functional stability has important implications for how we approach the relationship between biodiversity and ecosystem function. Just as links between declining species diversity and function may be less relevant than we thought (citations!) so too the realtionship between functional diversity and ecosystem function.  his is good news for ecosystem function! But we may also be measuring the wrong dimensions of the trait space or the wrong traits. Can't really say BEF needs to measure function directly because these communities aren't losing species or functional diversity - turnover is either within regular bounds or if not (if more turnover than normal) it's still maintaining community structure. And presuming that the original structure was good for function, maintaining it is also good for function! -->

@@

This study represents the best available data and biodiversity synthesis approaches.
Why then do we find no change in functional diversity despite extensive predictions to the contrary from both **theory** <!-- need more theory in the intro? -->and many local studies?

@@
